The Effects of the NBA COVID Bubble on the NBA Playoffs

### 2020-11-10 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

Comparisons between 2020 and 2017-19 home and away teams were made on home team
winning percentage, total team scoring and two-point, three-point and free throw
shooting. Comparing the differences in these metrics for home and away teams in
2020 vs previous years will provide valuable insights to the understanding of 
home-court advantage. We can see how going on the road may negatively impact 
away performance and how playing at home may positively impact home performance.
If there are differences in scoring for home or away teams the differences can
be used to show how home-court advantage affects overall performance of home
and away teams. While testing for differences in shooting will provide added
context for how home-court advantage specifically affects performance. 
Shooting percentages aren’t the only possible metrics affected by home-court
advantage, but they are the most obvious and certainly an important one. In
order to compare these differences simple two-sample z-tests were conducted.
Test for differences in means was conducted for comparing scoring totals, shown in 
Equation \@ref(eq:zmean). Tests for differences in proportions were conducted for
comparing winning percentage and the three different shooting percentages, shown
in Equation \@ref(eq:zprop). All 9 tests are conducted under the assumption and null
hypothesis that there should be no difference in statistics from 2020 compared
to those from 2017-2019.

\begin{align}
    z = \frac{(\bar{X}_1 - \bar{X}_2)-0}{\sqrt{\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2}}}.
    (\#eq:zmean)
\end{align}

\begin{align}
    z = \frac{(\hat{p}_1 - \hat{p}_2)-0}{\sqrt{\hat{p}(1-\hat{p})(\frac{1}{n_1}+\frac{1}{n_2})}}.
    (\#eq:zprop)
\end{align}

>>>>

Comparisons between 2020 and 2017-19 home and away teams were made on home team
winning percentage, total team scoring and two-point, three-point and free throw
shooting. Comparing the differences in these metrics for home and away teams in
2020 vs previous years will provide valuable insights to the understanding of 
home-court advantage. We can see how going on the road may negatively impact 
away performance and how playing at home may positively impact home performance.
If there are differences in scoring for home or away teams the differences can
be used to show how home-court advantage affects overall performance of home
and away teams. While testing for differences in shooting will provide added
context for how home-court advantage specifically affects performance. 
Shooting percentages aren’t the only possible metrics affected by home-court
advantage, but they are the most obvious and certainly an important one. In
order to compare these differences simple two-sample z-tests were conducted.

A test for differences in means was conducted to compare scoring totals. Let
$\bar{X}_1$ represent the average scoring total for 2020 and $\bar{X}_2$ represent
average scoring total over 2017-19. Similarly, $s_1^2$ represent the sample variance
of scoring for 2020 and $s_2^2$ represent sample variance of scoring over 2017-19.
Finally, let $n_1$ represent the number of games in 2020 and $n_2$ represent the number
of games from 2017-2019. Then the test statistic is constructed using Equation \@ref(eq:zmean).
\begin{align}
    z = \frac{(\bar{X}_1 - \bar{X}_2)-0}{\sqrt{\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2}}}.
    (\#eq:zmean)
\end{align}
We will test the null hypothesis that $\mu_1$=$\mu_2$. The test will be conducted twice
once comparing home scoring in 2020 to 2017-19 and once comparing away
scoring in 2020 to 2017-2019.

Tests for differences in proportions were conducted for comparing winning percentage
and the three different shooting percentages in 2020 to those of 2017-2019. Let
$\hat{p}_1$ represents sample proportion for 2020 and $\hat{p}_2$ represents sample
proportion for 2017-2019. While, $\hat{p}$ is $\hat{p}_1$ and $\hat{p}_2$ pooled 
together.Let $n_1$ represent number of observations for 2020 and $n_2$ represent 
number of observations for 2017-2019. Then the test statistic is constructed using
Equation \@ref(eq:zprop).
\begin{align}
    z = \frac{(\hat{p}_1 - \hat{p}_2)-0}{\sqrt{\hat{p}(1-\hat{p})(\frac{1}{n_1}+\frac{1}{n_2})}}.
    (\#eq:zprop)
\end{align}
We will test the null hypothesis that $p_1$=$p_2$. The test will be conducted seven times,
once comparing home winning percentage for 2020 to that of 2017-2019 and once for each of
two-point shooting, three-point shooting and free throws for both home and away teams
in 2020 vs. 2017-2019.
